### 2024-11-28 manuscript.tex
Addressing comments

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage{microtype}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
\setmainfont{EB Garamond}
\setmonofont[Scale=0.8]{SpaceMono}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald, Simon P. Alexander, Greg M. McFarquhar, John J. Cassano, Graeme Plank, Sean Hartery, Simon Parsons, Sally Garrett, and Alex J. Schuddeboom}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{13pt}{4pt}
\parskip=6pt
\parindent=0pt

@@

\begin{center} \Large \sffamily\textbf{Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses}\\[0.4cm]
\normalsize Peter Kuma$^\mathsf{1,2,\star}$, Frida A.-M. Bender$^\mathsf{1,2}$, Adrian J. McDonald$^\mathsf{3}$, Simon P. Alexander$^\mathsf{4,5}$, Greg M. McFarquhar$^\mathsf{6}$, John J. Cassano$^\mathsf{7,8,9}$, Graeme Plank$^\mathsf{3}$, Sean Hartery$^\mathsf{10}$, Simon Parsons$^\mathsf{11}$, Sally Garrett$^\mathsf{12}$, and Alex J. Schuddeboom$^\mathsf{3}$\\[0.4cm]
\footnotesize
$^\mathsf{1}$Department of Meteorology (MISU), Stockholm University, Stockholm,, Sweden\\
$^\mathsf{2}$Bolin Centre for Climate Research, Stockholm University, Stockholm, Sweden\\
$^\mathsf{3}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, Aotearoa/New Zealand\\
$^\mathsf{4}$Australian Antarctic Division, Kingston, Tasmania, Australia\\
$^\mathsf{5}$Australian Antarctic Program Partnership, Institute for Marine and Antarctic Studies, University of Tasmania, Hobart, Tasmania, Australia\\
$^\mathsf{6}$Cooperative Institute for Mesoscale Meteorological Studies, and School of Meteorology, University of Oklahoma, Norman, Oklahoma, USA\\
$^\mathsf{7}$Department of Atmospheric and Oceanic Sciences, University of Colorado Boulder, Boulder, Colorado, USA\\
$^\mathsf{8}$National Snow and Ice Data Center, University of Colorado Boulder, Boulder, Colorado, USA\\
$^\mathsf{9}$Department of Atmospheric and Oceanic Sciences, University of Colorado Boulder, Boulder, Colorado, USA\\[0.2cm]
$^\mathsf{10}$Department of Physics \& Atmospheric Science, Dalhousie University, Halifax, Canada\\
$^\mathsf{11}$New South Wales Department of Planning and Environment, Sydney, New South Wales, Australia\\
$^\mathsf{12}$New Zealand Defence Force, Wellington, New Zealand\\
$^\star$Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\normalsize \today\\[0.4cm]

@@

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is the 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). Due to the high resolution, parameterizations of convection and clouds
are avoided. Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Unstable conditions are the most problematic for the
models. In daily cloud cover, ICON and the reanalyses tend to be about 1 and 2
oktas clearer, respectively. Compared to radiosondes, potential temperature is
accurate in all, but ICON is too unstable over the low-latitude SO and too
humid in the boundary layer. MERRA-2 is too humid at all heights. SO cloud
biases are a substantial issue in the GSRM, but are an improvement over the
lower-resolution reanalyses. Explicitly resolved convection and cloud processes
were not enough to address the model cloud biases.

@@

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a trade-off with increasing parameterization complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advance in climate modelling.  Global storm-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models or global convection-permitting/-resolving
models are also sometimes used interchangeably with GSRMs but imply that clouds
or convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
over two 40-day time periods in summer (1 August -- 10 September 2016) and winter
(20 January -- 1 March 2020). A new one-year GSRM intercomparison is currently
proposed by \cite{takasuka2024}, with the hope of also evaluating the seasonal
cycle and large-scale circulation.  An alternative to using a computationally
costly GSRM is to train an artificial neural network on GSRM output and use it
for subgrid-scale clouds, as done with the GSRM ICON by \cite{grundner2022} and
\cite{grundner2023}.

@@

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW)
radiation. To first order, the total cloud fraction, cloud phase, and the
liquid and ice water path are the most important cloud properties influencing
SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. Second order effects on SW and LW radiation are associated
with the cloud droplet size distribution, ice crystal habit, cloud lifetime,
and direct radiative interaction with aerosols.  In the 6$^\mathrm{th}$ phase
of the Coupled Model Intercomparison Project [CMIP6; \cite{eyring2016}], the
cloud feedback has increased relative to CMIP5 \citep{zelinka2020}, this being
one of the main reasons for the higher climate sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and
being a lower priority region for numerical weather prediction (NWP) and
climate model development because of its distance from populated areas of
interest.  Nevertheless, radiation biases and changes over an area of its size
have a very substantial influence on the global climate, and the SO is an
important part of the global ocean conveyor belt.  Marine clouds have
a disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The relative longitudinal symmetry
of the SO means that model cloud biases tend to be similar across longitudes.
Here, we conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas over
the low-latitude SO it tends to be more neutral or negative in some models. Too
much absorbed SW radiation over the SO was also identified in the GSRM SCREAM
\cite{caldwell2021}. Compensating biases are possible, such as the `too few too
bright' cloud bias, characterised by too small cloud fraction and too large
cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a reasonable SW radiation balance by
reflecting too much SW radiation from clouds, but has too little cloud
area overall. A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging
Spectroradiometer (MODIS) data] and overestimates downwelling TOA SW radiation on
by approximately 10 Wm$^\mathrm{-2}$ in  [relative to the Clouds and the Earth’s
Radiant Energy System (CERES) data] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found the ICON cloud optical
thickness was underestimated relative to Himawari‐8 satellite observations, but
also identified large differences in cloud top phase.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/map_fig.pdf}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern},
RSV \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer},
and HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010--2021
(inclusive).  The dotted and dashed lines at 40°S and 55°S delineate the
Southern Ocean area of our analysis and its partitioning into two subsets,
respectively.  A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke Mehrtens,
Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma, University of Canterbury), \textbf{(d)} Vaisala
CL51 installation at the Macquarie Island station (© Jeff Aquilina, Bureau of
Meteorology).
% TODO: Ask for photo permission and info about license.
}
\label{fig:map}
\end{figure}

@@

Missing days in the ceilometer data were HMNZSW16 (7 days): 24--27 November, 10
December, 16--17 December 2016; Measurements of Aerosols, Radiation, and CloUds
over the Southern Ocean (MARCUS;  3 days): 8, 10 November, 10 December;
Macquarie Island Cloud Radiation Experiment (MICRE; 9 days): 7--8, 29 June, 5,
16 July, 15 August, 17 October 2016, 11 February, 21 March 2017; TAN1502 (1
day): 24 January.

@@

A coupled (atmosphere--ocean) GSRM version of the ICON model is in development
at the nextGEMS project \citep{hohenegger2023}. ICON is a very versatile model,
allowing for simulations ranging from coarse-resolution ESM simulations, GSRM
simulations, limited area model simulations, to large eddy simulations (LES),
for both weather prediction and climate projections. ICON uses the atmospheric
component ICON-A \citep{giorgetta2018}, whose physics is derived from ECHAM6
\citep{stevens2013}, and the ocean component ICON-O \citep{korn2022}. Earlier
runs of the GSRM ICON from DYAMOND were evaluated by \cite{mauritsen2022}.

@@

Because the model is free-running, weather and climate oscillations (such as
the El Niño--Southern Oscillation) are not expected to be equivalent to reality
at the same time and place. To compare with the observations collected on
different years (2010--2021, inclusive), we compared the model output with
observations at the same time of year and geographical location, as determined
for each data point such as a lidar profile or a radiosonde launch.

@@

TOA radiation quantities are taken from the CERES instruments on board of the
Terra and Aqua satellites \citep{wielicki1996,loeb2018}. In our analysis we
used the adjusted all sky SW and LW upwelling fluxes at TOA from the synoptic
TOA and surface fluxes and clouds 1 degree daily edition 4A product
(CER\_SYN1deg-Day\_Terra-Aqua-MODIS\_Edition4A)
\citep{doelling2013,doelling2016}.

@@

Precipitation can cause strong enough lidar backscattering to be recognised as
clouds by the threshold-based cloud detection method used in the ALCF. This is
undesirable if equivalent precipitation backscatter is not included in the
simulated lidar profiles. It was not possible to include precipitation
simulation in the ALCF due to the absence of required fields in the ICON model
output and the reanalysis data (the liquid and ice precipitation mass mixing
ratios) and because the required radiation calculations for precipitation are
currently not implemented in the ALCF. In order to achieve a fair comparison of
observations with models, one should exclude observed and simulated lidar
profiles with precipitation either manually or using an automated method. It is
relatively difficult to distinguish precipitation backscatter from cloud
backscatter in lidar observations, especially when only one wavelength channel
and no polarised channel are available. In models, the same can be accomplished
relatively easily by excluding profiles exceeding a certain amount of surface
precipitation flux. In the observations, using precipitation flux measurements
from rain gauges can be very unreliable on ships due to ship movement and
turbulence caused by nearby ship structures. Our analysis of rain gauge data
from the RV \emph{Tangaroa} showed large discrepancies between the rain gauge
time series and human-performed synoptic observations, as well as large
inconsistencies in the rain gauge time series. Human-performed observations of
precipitation presence or absence are expected to be reliable but only cover a
limited set of time instants. Therefore, it was desirable to implement a method
of detecting precipitation from observed backscatter profiles alone.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relevant for the subsequent analysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a shows a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied on the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, differences in the strength of the local
extremes of occurrence are notable, especially over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 years) and the fact that the model is free-running.

Fig. \ref{fig:cyclone-stability}b shows a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}). Relatively unstable conditions are prevalent
in the middle SO (50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. Relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 reanalysis is also substantially more stable than
ICON across the whole region.

>>>>

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage{microtype}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
\setmainfont{EB Garamond}
\setmonofont[Scale=0.8]{SpaceMono}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald, Simon P. Alexander, Greg M. McFarquhar, John J. Cassano, Graeme E. Plank, Sean Hartery, Simon Parsons, Sally Garrett, and Alex J. Schuddeboom}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{13pt}{4pt}
\parskip=6pt
\parindent=0pt

@@

\begin{center} \Large \sffamily\textbf{Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses}\\[0.4cm]
\normalsize Peter Kuma$^\mathsf{1,2,\star}$, Frida A.-M. Bender$^\mathsf{1,2}$, Adrian J. McDonald$^\mathsf{3}$, Simon P. Alexander$^\mathsf{4,5}$, Greg M. McFarquhar$^\mathsf{6,7}$, John J. Cassano$^\mathsf{8,9,10}$, Graeme E. Plank$^\mathsf{3}$, Sean Hartery$^\mathsf{11}$, Simon Parsons$^\mathsf{12}$, Sally Garrett$^\mathsf{13}$, and Alex J. Schuddeboom$^\mathsf{3}$\\[0.4cm]
\footnotesize
$^\mathsf{1}$Department of Meteorology (MISU), Stockholm University, Stockholm, Sweden\\
$^\mathsf{2}$Bolin Centre for Climate Research, Stockholm University, Stockholm, Sweden\\
$^\mathsf{3}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, Aotearoa/New Zealand\\
$^\mathsf{4}$Australian Antarctic Division, Kingston, Tasmania, Australia\\
$^\mathsf{5}$Australian Antarctic Program Partnership, Institute for Marine and Antarctic Studies, University of Tasmania, Hobart, Tasmania, Australia\\
$^\mathsf{6}$Cooperative Institute of Severe and High Impact Weather Research and Operations, University of Oklahoma, Norman, OK, USA\\
$^\mathsf{7}$School of Meteorology, University of Oklahoma, Norman, OK, USA\\
$^\mathsf{8}$Cooperative Institute for Research in Environmental Sciences, University of Colorado, Boulder, CO, USA\\
$^\mathsf{9}$National Snow and Ice Data Center, University of Colorado, Boulder, CO, USA\\
$^\mathsf{10}$Department of Atmospheric and Oceanic Sciences, University of Colorado, Boulder, CO, USA\\[0.2cm]
$^\mathsf{11}$Department of Physics \& Atmospheric Science, Dalhousie University, Halifax, Canada\\
$^\mathsf{12}$New South Wales Department of Planning and Environment, Sydney, New South Wales, Australia\\
$^\mathsf{13}$New Zealand Defence Force, Wellington, New Zealand\\
$^\star$Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\normalsize \today\\[0.4cm]

@@

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is the 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). The high resolution allows for parameterizations of convection and
clouds to be avoided. Standard-resolution models have substantial cloud biases
over the Southern Ocean (SO), affecting radiation and sea surface temperature.
We evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Less stable conditions are the most problematic for
ICON and the reanalyses.  In daily cloud cover, ICON and the reanalyses tend to
be about 1 and 2 oktas clearer, respectively. Compared to radiosondes,
potential temperature is accurate in the reanalyses, but ICON underestimates
stability over the low-latitude SO and too humid in the boundary layer. MERRA-2
is too humid at all heights. SO cloud biases remain a substantial issue in the
GSRM, but are an improvement over the lower-resolution reanalyses. Explicitly
resolved convection and cloud processes were not enough to address the model
cloud biases.

@@

Increasing climate model resolution is one way of improving model accuracy of
representation of the climate system \citep{mauritsen2022}. It has been
practiced since the advent of climate modelling as more computational power,
memory, and storage capacity become available. It is, however, often not as
easy as changing the grid size because of the complex interplay between model
dynamics and physics, which necessitates adjusting and tuning all components
together.  Increasing resolution is of course limited by the available
computational power and a trade-off with increasing parameterization
complexity, which is another way of improving model accuracy.  Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPUs) is progressing to enable km-scale (also called
k-scale) Earth system models (ESMs) and coupled atmosphere--ocean general
circulation models (AOGCMs) in research conditions today and operationally in
the forthcoming years.  Therefore, it represents a natural advance in climate
modelling.  Global storm-resolving models (GSRMs) are emerging as a new front
in the development of high-resolution global climate models, with horizontal
grid resolutions of about 2--8 km \citep{satoh2019,stevens2019}. This is enough
to resolve mesoscale convective storms, but smaller-scale convective plumes and
cloud structure remain unresolved. At an approximately 5-km scale,
non-hydrostatic processes also become important \citep{weisman1997}, and for
this reason such models are generally non-hydrostatic. The terms global
cloud-resolving models or global convection-permitting/-resolving models are
also sometimes used interchangeably with GSRMs but imply that clouds or
convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
over two 40-day time periods in summer (1 August -- 10 September 2016) and
winter (20 January -- 1 March 2020). A new one-year GSRM intercomparison is
currently proposed by \cite{takasuka2024}, with the hope of also evaluating the
seasonal cycle and large-scale circulation.  An alternative to using a
computationally costly GSRM is to train an artificial neural network on GSRM
output and use it for subgrid-scale clouds, as done with the GSRM ICON by
\cite{grundner2022} and \cite{grundner2023}.

@@

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW)
radiation. To first order, the total cloud fraction, cloud phase, and the
liquid and ice water path are the most important cloud properties influencing
SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. Second order effects on SW and LW radiation are associated
with the cloud droplet size distribution, ice crystal habit, cloud lifetime,
and direct radiative interaction with aerosols.  In the 6$^\mathrm{th}$ phase
of the Coupled Model Intercomparison Project [CMIP6; \cite{eyring2016}], the
cloud feedback has increased relative to CMIP5 \citep{zelinka2020}, which is
one of the main reasons for the higher climate sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and being a lower
priority region for numerical weather prediction (NWP) and climate model
development because of its distance from populated areas.  Nevertheless,
radiation biases and changes over an area of its size have a substantial
influence on the global climate, and the SO is an important part of the global
ocean conveyor belt.  Marine clouds have a disproportionate effect on top of
atmosphere (TOA) SW radiation due to the relatively low albedo of the sea
surface.  The relative longitudinal symmetry of the SO means that model cloud
biases tend to be similar across longitudes.  Here, we conventionally refer to
the SO as ocean regions south of 40°S, low-latitude SO as 40--55°S and
high-latitude SO as south of 55°S.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas
over the low-latitude SO it tends to be more neutral or negative in some
models. Too much absorbed SW radiation over the SO was also identified in the
GSRM SCREAM \cite{caldwell2021}. Compensating biases are possible, such as the
`too few too bright' cloud bias, characterised by too small cloud fraction and
too large cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a reasonable SW radiation balance by
reflecting too much SW radiation from clouds, but has too little cloud area
overall. A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging Spectroradiometer
(MODIS) data] and overestimates downwelling TOA SW radiation on by
approximately 10 Wm$^\mathrm{-2}$ [relative to the Clouds and the Earth’s
Radiant Energy System (CERES) data] in the highest model resolution run (2.5
km).  \cite{zhao2022} reported a similar SW radiation bias in five analysed
CMIP6 models over high-latitude SO and the total cloud fraction underestimation
on the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found
the ICON cloud optical thickness was underestimated relative to Himawari‐8
satellite observations, but also identified large differences in cloud top
phase.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/map_fig.pdf}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern},
RSV \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer},
and HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010--2021
(inclusive).  The dotted and dashed lines at 40°S and 55°S delineate the
Southern Ocean area of our analysis and its partitioning into two subsets,
respectively.  A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke Mehrtens,
Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma, University of Canterbury), \textbf{(d)} Vaisala
CL51 (© Jeff Aquilina, Bureau of Meteorology).
}
\label{fig:map}
\end{figure}

@@

Missing days in the ceilometer data were HMNZSW16 (7 days): 24--27 November, 10
December, 16--17 December 2016; Measurements of Aerosols, Radiation, and CloUds
over the Southern Ocean (MARCUS;  3 days): 8, 10 November, 10 December 2017;
Macquarie Island Cloud Radiation Experiment (MICRE; 9 days): 7--8, 29 June, 5,
16 July, 15 August, 17 October 2016, 11 February, 21 March 2017; TAN1502 (1
day): 24 January.

@@

A coupled (atmosphere--ocean) GSRM version of the ICON model is in development
at the nextGEMS project \citep{hohenegger2023}. ICON is an exceptionally
versatile model, allowing for simulations ranging from coarse-resolution ESM
simulations, GSRM simulations, limited area model simulations, to large eddy
simulations (LES), for both weather prediction and climate projections. ICON
uses the atmospheric component ICON-A \citep{giorgetta2018}, whose physics is
derived from ECHAM6 \citep{stevens2013}, and the ocean component ICON-O
\citep{korn2022}. Earlier runs of the GSRM ICON from DYAMOND were evaluated by
\cite{mauritsen2022}.

@@

Because the model is free-running, weather and climate oscillations (such as
the El Niño--Southern Oscillation) are not expected to be equivalent to reality
at the same time and place. To compare with the observations collected in
different years (2010--2021, inclusive), we compared the model output with
observations at the same time of year and geographical location, as determined
for each data point such as a lidar profile or a radiosonde launch.

@@

TOA radiation quantities are taken from the CERES instruments on board the
Terra and Aqua satellites \citep{wielicki1996,loeb2018}. In our analysis we
used the adjusted all sky SW and LW upwelling fluxes at TOA from the synoptic
TOA and surface fluxes and clouds 1 degree daily edition 4A product
(CER\_SYN1deg-Day\_Terra-Aqua-MODIS\_Edition4A)
\citep{doelling2013,doelling2016}.

@@

Precipitation can cause strong enough lidar backscattering to be recognised as
clouds by the threshold-based cloud detection method used in the ALCF. This is
undesirable if equivalent precipitation backscatter is not included in the
simulated lidar profiles. It was not possible to include precipitation
simulation in the ALCF due to the absence of required fields in the ICON model
output and the reanalysis data (the liquid and ice precipitation mass mixing
ratios). The required radiation calculations for precipitation are also
currently not implemented in the ALCF, even though this is a planned feature.
In order to achieve a fair comparison of observations with models, one should
exclude observed and simulated lidar profiles with precipitation either
manually or using an automated method. It is relatively difficult to
distinguish precipitation backscatter from cloud backscatter in lidar
observations, especially when only one wavelength channel and no polarised
channel are available. In models, the same can be accomplished relatively
easily by excluding profiles exceeding a certain amount of surface
precipitation flux. In the observations, using precipitation flux measurements
from rain gauges can be very unreliable on ships due to ship movement,
turbulence caused by nearby ship structures, and sea spray. Our analysis of
rain gauge data from the RV \emph{Tangaroa} showed large discrepancies between
the rain gauge time series and human-performed synoptic observations, as well
as large inconsistencies in the rain gauge time series. Human-performed
observations of precipitation presence or absence are expected to be reliable
but only cover a limited set of time instants. Therefore, it was desirable to
implement a method of detecting precipitation from observed backscatter
profiles alone.

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relevant for the subsequent analysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a, b show a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied on the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, differences in the strength of the local
extremes of occurrence are notable, especially over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adare, which is more cyclonic
in ICON. These differences might, however, stem from the relatively short time
periods of comparison (4 years) and the fact that the model is free-running.

Fig. \ref{fig:cyclone-stability}c, d show a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}). Relatively unstable conditions are prevalent
in the middle SO (50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. Relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 reanalysis is also substantially more stable than
ICON across the whole region.
